---
title: "lab_0"
author: "Kevin Luo"
date: "2019/9/1"
output: html_document
---

```{r setup, include=FALSE}
library(tidyr)
library(tidyverse)
library(dplyr)
```

Part 1) Basic Vector and Matrix manipulations in R
1.1-1.3 Generating matrices in R
```{r}
x <- 1:9
matrix(x, nrow = 3, ncol = 3)
matrix(x, nrow = 3, byrow = TRUE)
diag(nrow = 5)
```

1.4-1.5 Combining Matrices
```{r}
a1 <-c(2, 3, 6, 7, 10)
a2 <-c(1.88, 2.05, 1.70, 1.60, 1.78)
a3 <-c(80, 90, 70, 50, 75)
A <- matrix(c(a1,a2,a3), nrow = 5)

b1 <-c(1, 4, 5, 8, 9)
b2 <-c(1.22, 1.05, 3.60, 0.40, 2.54)
b3 <-c(20, 40, 30, 80, 100)
B <- matrix(c(b1,b2,b3), nrow = 3, byrow = TRUE)
```

1.6 Calculate Products using %*%
```{r}
A %*% B
B %*% A
t(A) %*% t(B)
t(B) %*% t(A)
```

1.7 obtain a linear function using matrix multiplication:
```{r}
matrix_data <- iris %>%
  select(1:4) %>%
  pull() %>%
  matrix(ncol = 4)
eff <- 1:4
temp <- eff %*% t(matrix_data)
```

1.8-1.9 write a vnorm function & normalization
```{r}
vnorm <- function(a){
  return(sqrt(t(a) %*% a))
}
# testing it with 1:5
v <- 1:5
vnorm(v)
u <- v/c(vnorm(v))
```

1.10 checking a square matrix
```{r}
is_square <- function(A){
  if (nrow(A) == ncol(A)){
    return(TRUE)
  }else{
    return(FALSE)
  }
}
```

1.11 calculate the trace
```{r}
mtrace <- function(A){
  if (is_square(A) == FALSE){
    return("The input is not a square matrix")
  }else{
    tr = 0
    for (i in 1:ncol(A)) {
      tr = tr + A[i,i]
    }
    return(tr)
  }
}
```

1.12 Some verifications
```{r}
A <- matrix(1:25, nrow = 5)
B <- diag(nrow = 5)
C <- 5
# You can arbitrarily pick the values of A,B and C, using the verification function will give you a return value of TRUE
verification <- function(A,B,C){
  if(mtrace(A+B)-mtrace(A)-mtrace(B) == 0 &
     mtrace(C*A)-C*mtrace(A) == 0 &
     mtrace(t(A)*B) == mtrace(A*t(B)) &
     mtrace(B*t(A)) == mtrace(A*t(B)) &
     mtrace(t(B)*A) == mtrace(A*t(B))){
    return(TRUE)
  }else{
    return(FALSE)
  }
}
verification(A,B,C)
```


Part 2) Transformation and Scaling Operations
2.1-2.7 Creating a matrix, mean-centering, maximum calculating
```{r}
M <- mtcars %>%
  select(c("mpg", "disp", "hp", "drat", "wt")) %>%
  as.matrix()

apply(M, 2, mean)

Mc <- scale(M, scale = FALSE)
apply(Mc, 2, mean)

temp <- sweep(M, 2, apply(M,2,mean))
apply(temp, 2, mean)
# We can use (Mc-temp) to verify that we're getting the same results
head(Mc-temp)
# We'll get a matrix of zero
apply(M,2,max)
```

2.8 0-1 Scaling
```{r}
M_scale <- apply(M, 2, FUN = function(X){(X-min(X))/(max(X)-min(X))})
```

2.9 Calculating the covariance matrix
```{r}
meanM <- apply(M, 2, mean)
one <- rep(1,nrow(M))

COV <- (1/(nrow(M)-1))*(t(M) - (meanM) %*% t(one)) %*% t(t(M) - (meanM) %*% t(one))
```

2.10 Calculating the correlation matrix
```{r}
temp <- COV
cor <- matrix(0, nrow = 5, ncol = 5)
cor
for (i in 1:5) {
  for (j in 1:5){
    cor[i,j] = temp[i,j] / sqrt(temp[i,i]*temp[j,j])
  }
}
```

2.11 return a dummy matrix
```{r}
dummify <- function(data, all=TRUE){
  if (all == TRUE){
    name <- unique(data)
    dummy <- matrix(0, nrow = length(data), ncol = length(name))
    for (i in 1:length(data)){
      for (j in 1:length(name)){
        if (data[i] == name[j]){
          dummy[i,j] = 1
        }
      }
    }
  }else{
    name <- unique(data)
    name <- name[-length(name)]
    dummy <- matrix(0, nrow = length(data), ncol = length(name))
    for (i in 1:length(data)){
      for (j in 1:length(name)){
        if (data[i] == name[j]){
          dummy[i,j] = 1
        }
      }
    }
  }
  return(dummy)
}

# for test purpose
cyl <- factor(mtcars$cyl)
# all categories
CYL1 <- dummify(cyl, all = TRUE) # minus one category
CYL2 <- dummify(cyl, all = FALSE)

```

3.11 Calculating a crosstable
```{r}
crosstable <- function(data1, data2){
  return(t(dummify(data1)) %*% dummify(data2))
}
# For testing purpose:
cyl <- factor(mtcars$cyl)
gear <- factor(mtcars$gear)
xtb <- crosstable(cyl, gear)
# Note that the output matrix differs from the standard table only in the sequence of columns and rows. They're identical when we examine the pairs of values.
```

