---
title: "Assignment 8"
author: "Kaicheng Luo"
date: "2019/11/3"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, message=F, warning=F}
library(tidyr)
library(tidyverse)
library(ggplot2)
library(dplyr)
data(iris)
```

```{r}
data <- iris
tss <- function(x){
  return(sum((x - mean(x))^2))
}
tss(iris$Sepal.Length)
```

```{r}
bss <- function(x, group){
  if (length(x) != length(group)){
    stop()
  }
  group = factor(group)
  result = 0
  m <- mean(x)
  for (i in levels(group)){
    subgroup <- x[group == i]
    result = result + length(subgroup) * (mean(subgroup) - m)^2
  }
  return(result)
}
bss(iris$Sepal.Length, iris$Species)
```

```{r}
wss <- function(x, group){
  if (length(x) != length(group)){
    stop()
  }
  group = factor(group)
  result = 0
  for (i in levels(group)){
    subgroup <- x[group == i]
    result = result + tss(subgroup)
  }
  return(result)
}
wss(iris$Sepal.Length, iris$Species)
# Note that by decomposition, this shall return zero
tss(iris$Sepal.Length) - wss(iris$Sepal.Length, iris$Species) - bss(iris$Sepal.Length, iris$Species)
```

```{r}
cor_ratio <- function(x, group){
  return(bss(x, group) / tss(x))
}
cor_ratio(iris$Sepal.Length, iris$Species)

F_ratio <- function(x, group){
  return(
    bss(x, group) * (length(x) - length(levels(group))) / wss(x, group) * (length(levels(group)) - 1)
  )
}
F_ratio(iris$Sepal.Length, iris$Species)
```

```{r}
list("Slength" = cor_ratio(iris$Sepal.Length, iris$Species), "Swidth" = cor_ratio(iris$Sepal.Width, iris$Species), "Plength" = cor_ratio(iris$Petal.Length, iris$Species), "Pwidth" = cor_ratio(iris$Petal.Width, iris$Species))
data.frame("Predictors" = c("Slength","Swidth","Plength","Pwidth"), "Cor" = c(cor_ratio(iris$Sepal.Length, iris$Species), cor_ratio(iris$Sepal.Width, iris$Species), cor_ratio(iris$Petal.Length, iris$Species), cor_ratio(iris$Petal.Width, iris$Species))) %>% arrange(by = Cor)

data.frame("Predictors" = c("Slength","Swidth","Plength","Pwidth"), "Cor" = c(cor_ratio(iris$Sepal.Length, iris$Species), cor_ratio(iris$Sepal.Width, iris$Species), cor_ratio(iris$Petal.Length, iris$Species), cor_ratio(iris$Petal.Width, iris$Species))) %>% arrange(by = Cor) %>%
  ggplot() + theme_bw() +
    geom_bar(aes(x = Predictors, y = Cor), stat = "identity", width = .6)
```

```{r}
list("Slength" = F_ratio(iris$Sepal.Length, iris$Species), "Swidth" = F_ratio(iris$Sepal.Width, iris$Species), "Plength" = F_ratio(iris$Petal.Length, iris$Species), "Pwidth" = F_ratio(iris$Petal.Width, iris$Species))
data.frame("Predictors" = c("Slength","Swidth","Plength","Pwidth"), "F_ratio" = c(F_ratio(iris$Sepal.Length, iris$Species), F_ratio(iris$Sepal.Width, iris$Species), F_ratio(iris$Petal.Length, iris$Species), F_ratio(iris$Petal.Width, iris$Species))) %>% arrange(by = F_ratio)

data.frame("Predictors" = c("Slength","Swidth","Plength","Pwidth"), "F_ratio" = c(F_ratio(iris$Sepal.Length, iris$Species), F_ratio(iris$Sepal.Width, iris$Species), F_ratio(iris$Petal.Length, iris$Species), F_ratio(iris$Petal.Width, iris$Species))) %>% arrange(by = F_ratio) %>%
  ggplot() + theme_bw() +
    geom_bar(aes(x = Predictors, y = F_ratio), stat = "identity", width = .6)
```

```{r}
total_variance <- function(x){
  X <- as.matrix(x - (matrix(1, nrow = nrow(x)) %*% t(apply(x, 2, mean))))
  return(t(X) %*% X / (nrow(X)-1))
}
total_variance(iris[,1:4])
# Compare with the built-in functions, they are identical!
var(iris[,1:4])
```

```{r}
between_variance <- function(x, group){
  m <- apply(x, 2, mean)
  group <- factor(group)
  result = matrix(0, nrow = ncol(x), ncol = ncol(x))
  for (i in levels(group)){
    subgroup = x[group == i, ]
    result = result + nrow(subgroup) / (nrow(x)-1) *(apply(subgroup, 2, mean) - m) %*% t(apply(subgroup, 2, mean) - m)
  }
  return(result)
}
between_variance(iris[,1:4], iris[,5])
```

```{r}
within_variance <- function(x, group){
  group <- factor(group)
  result = matrix(0, nrow = ncol(x), ncol = ncol(x))
  for (i in levels(group)){
    subgroup = x[group == i,]
    result = result + (nrow(subgroup)-1) * total_variance(subgroup) / (nrow(x) - 1)
  }
  return(result)
}
within_variance(iris[ ,1:4], iris[ ,5])

# Here we can verify our decompostion formula
round(total_variance(iris[,1:4]) - within_variance(iris[ ,1:4], iris[ ,5]) - between_variance(iris[,1:4], iris[,5]))
```

```{r}
# confirm V = B + W
Viris <- total_variance(iris[ ,1:4]) 
Viris
#B+W
Biris <- between_variance(iris[ ,1:4], iris$Species) 
Wiris <- within_variance(iris[ ,1:4], iris$Species) 
Biris + Wiris
```

```{r}
C <- function(x, group){
  m <- apply(x, 2, mean)
  group <- factor(group)
  result = matrix(0, nrow = ncol(x), ncol = ncol(x))
  for (j in 1:ncol(x)){
    k = 0
    for (i in levels(group)){
      subgroup = x[group == i, ]
      k = k+1
      result[j,k] = sqrt(nrow(subgroup) / (nrow(x)-1)) * (mean(subgroup[,j]) - m[j])
    }
  }
  return(result)
}
# Verify that the decompostion is correct
round(C(iris[ ,1:4], iris$Species) %*% t(C(iris[ ,1:4], iris$Species)) -between_variance(iris[ ,1:4], iris$Species))
C <- C(iris[ ,1:4], iris$Species)
```

```{r}
result <- eigen(t(C) %*% solve(Wiris) %*% C)
w <- result$vectors
u <- solve(Wiris) %*%C %*% w
z1 <- as.matrix(iris[,1:4]) %*% u[,1]
z2 <- as.matrix(iris[,1:4]) %*% u[,2]
plotdata <- data.frame("z1" = z1, "z2" = z2, group = iris$Species)
ggplot() +
  geom_point(aes(x = z1, y = z2, color = "setosa"), data = plotdata %>% filter(group == "setosa")) +
  geom_point(aes(x = z1, y = z2, color = "versicolor"), data = plotdata %>% filter(group == "versicolor")) +
  geom_point(aes(x = z1, y = z2, color = "virginica"), data = plotdata %>% filter(group == "virginica"))
```

```{r}
result <- eigen(t(as.matrix(iris[,1:4])) %*% as.matrix(iris[1:4]))
w <- result$vectors
z1 <- as.matrix(iris[,1:4]) %*% w[,1]
z2 <- as.matrix(iris[,1:4]) %*% w[,2]
plotdata <- data.frame("z1" = z1, "z2" = z2, group = iris$Species)
ggplot() +
  geom_point(aes(x = z1, y = z2, color = "setosa"), data = plotdata %>% filter(group == "setosa")) +
  geom_point(aes(x = z1, y = z2, color = "versicolor"), data = plotdata %>% filter(group == "versicolor")) +
  geom_point(aes(x = z1, y = z2, color = "virginica"), data = plotdata %>% filter(group == "virginica"))
```

The princical component projection offers us the largest total variance, regardless of the information of the group of variables. There're still significant overlapping in the axis of the first principal component that leads to a somehow difficult task to classify objects. CDA, on the other hand, offers us the best "classification" projection, taking the information of group into consideration.

```{r}
distance <-  function(x, newx, Wiris){
  return((x - newx) %*% solve(Wiris) %*% t(x-newx))
}
x <- iris[,1:4]
group <- iris$Species
Wiris <- within_variance(x, group)
group <- factor(group)
centroid <- matrix(0, nrow = ncol(x), ncol = length(levels(group)))
k = 0
for (i in levels(group)){
  k = k+1
  centroid[,k] = apply(x[group == i, ],2,mean)
}
x1 =c(5.0,3.0,1.5,0.5) 
x2 =c(5.5,3.0,6.0,2.0) 
x3 =c(6.0,3.0,4.0,1.0)
x4 =c(5.0,3.0,1.0,0.5)
```

```{r}
CDApred <- function(x, W = Wiris, centroid = centroid, group = iris$Species){
  dis = c()
  for (i in 1:length(levels(group))){
    dis = c(dis, distance(centroid[,i], t(x), W))
  }
  if (which.min(dis) == 1){print(paste("It shall be classified into setosa. The Mdistance is ", dis[1]))}
  if (which.min(dis) == 2){print(paste("It shall be classified into versicolor. The Mdistance is ", dis[2]))}
  if (which.min(dis) == 3){print(paste("It shall be classified into virginica. The Mdistance is ", dis[3]))}
}

CDApred(x1, Wiris, centroid)
CDApred(x2, Wiris, centroid)
CDApred(x3, Wiris, centroid)
CDApred(x4, Wiris, centroid)
```

